---
title: "dplyr"
subtitle: ""
author: "The R Bootcamp<br/>Twitter: <a href='https://twitter.com/therbootcamp'>@therbootcamp</a>"
date: "June 2017"
output:
  xaringan::moon_reader:
    css: ["default", "my-theme.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
# see: https://github.com/yihui/xaringan
# install.packages("xaringan")
# see: 
# https://github.com/yihui/xaringan/wiki
# https://github.com/gnab/remark/wiki/Markdown
options(width=110)
options(digits = 4)
```

# R

From [Wikipedia](https://en.wikipedia.org/wiki/Statistical_model) (emphasis added):

> R is an **open source programming language** and software environment for **statistical computing and graphics** that is supported by the R Foundation for Statistical Computing. The R language is **widely used among statisticians and data miners** for developing statistical software and data analysis. Polls, surveys of data miners, and studies of scholarly literature databases show that **R's popularity has increased substantially in recent years**.

> R is a GNU package. The source code for the R software environment is written primarily in **C, Fortran, and R**. R is freely available under the GNU General Public License, and pre-compiled binary versions are provided for various operating systems. While R has a command line interface, there are several **graphical front-ends available**.

---

# Programming language

From [Wikipedia](https://en.wikipedia.org/wiki/Statistical_model) (emphasis added):

> A programming language is a **formal language** that specifies a set of instructions that can be used to produce various kinds of output. Programming languages generally consist of **instructions for a computer**. Programming languages can be used to create programs that **implement specific algorithms**.


.pull-left4[
### Algorithm
```{block}
Load data
Extract variables
Run analysis
Print result
```
]
.pull-right4[
### Implementation in R
```{r, results='hide'}
data = read.table('my_data.txt')
variables = data[,c('group','variable')]
analysis = lm(variable ~ group, data = variables)
summary(analysis)
```
]

