---
title: "Statistics"
subtitle: ""
author: "The R Bootcamp<br/>Twitter: <a href='https://twitter.com/therbootcamp'>@therbootcamp</a>"
date: "September 2017"
output:
  xaringan::moon_reader:
    css: ["default", "my-theme.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
# see: https://github.com/yihui/xaringan
# install.packages("xaringan")
# see: 
# https://github.com/yihui/xaringan/wiki
# https://github.com/gnab/remark/wiki/Markdown
options(width=110)
options(digits = 4)
```

# Statistics
.pull-left3[

asdf

]


.pull-right3[

### Questions you can answer with dplyr

> *Can you calculate the mean survival times for each treatment separated by gender and time?*

> *I need to know the mean birth rate only for countries in Africa from 1980 to 1980.*

> *What percent of female patients had adverse events to drug X during weeks 5 through 10?*

]
---
# Two types of statistics: Descriptive and Inferential

.pull-left3[

### Descriptive

- Also called *sample statistics*
- Used to describe general characteristics of a sample
- Descriptive statistics typically a single scaler value


#### Examples

| Statistic| R code|
|------:|----:|
|     Mean|    `mean(x)`|
|     Median|    `median(x)`|
|     Mode|   `mode(x)`|
|     Standard Deviation|    `sd(x)`|
]

.pull-right3[
#### R implimentation

```{r}
sd(c(5, 3, 6, 3, 2, 6))  # Standard deviation

mean(ChickWeight$weight) # Mean weight

median(ChickWeight$Time) # Mean Time
```


]




---
# Two types of statistics: Descriptive and Inferential

.pull-left3[

### Inferential

- Used to make inferences about a larger population.
- Typically done in tandem with a *hypothesis test*
- Hypothesis tests typically return lists of named items

#### Examples

| Hypothesis Test| R code|
|------:|----:|
|     T-test|    `t.test()`|
|     Correlation Test|    `cor.test()`|
|     Chi-Square Test|   `chisq.test()`|
|     ANOVA|    `aov()`|
]

.pull-right3[
#### R implimentation

```{r}
t.test(x = c(4, 3, 6, 5, 3, 2),
       mu = 0,
       alternative = "two.sided")
```


]


---
# Inferential Statistics

.pull-left3[

### Assigning hypothesis test objects

- Most hypothesis tests return an object of class `"htest"` which contain many values
- You can assign the results of a hypothesis test to an object, and then extract the info you want

#### Examples of what's in htest objects

| Element| Result|
|------:|----:|
|    `x$statistic`|    A test statistic|
|     `x$parameter`|    Degrees of freedom|
|     `x$p.value`|   The p-value|
|     `x$conf.int`|  Confidence interval|
]

.pull-right3[
#### R implimentation

```{r}
# One-sample t-test comparing the weights of chicks to 120

weight.tt <- t.test(x = ChickWeight$weight,
                    mu = 120,
                    alternative = "two.sided")

class(weight.tt)

# What's in the weight.tt object?
names(weight.tt)

```


]



---
# Inferential Statistics

.pull-left3[

### Hyopthesis tests

- Different hypothesis tests require different inputs
- A simple one-sample t-test requires just a vector, while an ANOVA requires many inputs.

#### Examples

| Hypothesis Test| R code|
|------:|----:|
|     T-test|    `t.test()`|
|     Correlation Test|    `cor.test()`|
|     Chi-Square Test|   `chisq.test()`|
|     ANOVA|    `aov()`|
]

.pull-right3[
#### R implimentation

```{r}
t.test(x = c(4, 3, 6, 5, 3, 2),
       mu = 0,
       alternative = "two.sided")
```


]


---
# Inferential Statistics


.pull-left4[

> To see all the named elements from a hypothesis test object x, run `names(x)`:

> `names(x)`

> Then you can access the objects by name with $:

> `x$p.value`
> `x$statistic`
> `x$parameter`
> `x$conf.int`

]

.pull-right4[

```{r}
# One-sample t-test comparing the weights of chicks to 120

weight.tt <- t.test(x = ChickWeight$weight,
                    mu = 120,
                    alternative = "two.sided")

names(weight.tt)

weight.tt$statistic # What was the test statistic?
```
]


